---
title: "Bayesian Analysis of Aviation Accidents"
author: "Your Group Name"
date: today
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
  html:
    toc: true
    code-fold: true
    code-summary: "Show code"
execute:
  warning: false
  message: false
  cache: true
---

# Introduction

Brief introduction to your research question and motivation.

# Data Description

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(brms)
library(bayesplot)
library(loo)
library(posterior)
library(skimr)
library(knitr)
library(patchwork)

# Set plotting theme
theme_set(theme_minimal())

# Set options for better output
options(mc.cores = parallel::detectCores())
```

## Data Loading

```{r}
#| label: load-data

# Load your plane crash data
crashes <- read_csv("data/aircrashdata.csv")

# Clean column names for easier use
crashes <- crashes %>%
  rename(
    year = Year,
    quarter = Quarter,
    month = Month,
    day = Day,
    country = `Country/Region`,
    aircraft_manufacturer = `Aircraft Manufacturer`,
    aircraft_type = Aircraft,
    location = Location,
    operator = Operator,
    ground = Ground,
    fatalities = `Fatalities (air)`,
    aboard = Aboard
  )

# Display structure
glimpse(crashes)
```

## Data Summary

```{r}
#| label: data-summary

# Overview statistics
skim(crashes)

# Display first few rows
head(crashes) %>% 
  kable(caption = "First few rows of plane crash data")

# Print column names to verify
cat("Column names in dataset:\n")
names(crashes)
```

# Exploratory Data Analysis

## Response Variable Distribution

```{r}
#| label: fig-response-dist
#| fig-cap: "Distribution of fatalities in aviation accidents"
#| fig-width: 10
#| fig-height: 4

p1 <- ggplot(crashes, aes(x = fatalities)) +
  geom_histogram(bins = 50, fill = "steelblue", alpha = 0.7) +
  labs(title = "Histogram of Fatalities",
       x = "Number of Fatalities",
       y = "Count")

p2 <- ggplot(crashes, aes(x = fatalities)) +
  geom_histogram(bins = 50, fill = "steelblue", alpha = 0.7) +
  scale_y_log10() +
  labs(title = "Histogram (Log Scale)",
       x = "Number of Fatalities",
       y = "Count (log scale)")

p1 + p2
```

## Distributional Characteristics

```{r}
#| label: dist-stats

# Calculate key statistics
fatality_stats <- crashes %>%
  summarise(
    mean = mean(fatalities, na.rm = TRUE),
    variance = var(fatalities, na.rm = TRUE),
    var_mean_ratio = variance / mean,
    prop_zeros = mean(fatalities == 0, na.rm = TRUE),
    median = median(fatalities, na.rm = TRUE),
    max = max(fatalities, na.rm = TRUE),
    n = n()
  )

fatality_stats %>% 
  kable(digits = 2, caption = "Key statistics for fatalities")
```

The variance-to-mean ratio suggests whether we need Poisson or Negative Binomial distribution.

## Additional Variables

```{r}
#| label: additional-vars

# Create additional useful variables
crashes <- crashes %>%
  mutate(
    # Fatality rate
    fatality_rate = fatalities / aboard,
    # Total deaths (air + ground)
    total_deaths = fatalities + ground,
    # Binary: any fatalities
    any_fatalities = ifelse(fatalities > 0, 1, 0),
    # Decade
    decade = floor(year / 10) * 10
  )

# Summary of new variables
crashes %>%
  summarise(
    mean_fatality_rate = mean(fatality_rate, na.rm = TRUE),
    mean_total_deaths = mean(total_deaths, na.rm = TRUE),
    prop_with_fatalities = mean(any_fatalities, na.rm = TRUE)
  ) %>%
  kable(digits = 3, caption = "Additional derived variables")
```

## Comparing Theoretical Distributions

```{r}
#| label: fig-compare-dists
#| fig-cap: "Comparison of observed data with theoretical distributions"
#| fig-width: 10
#| fig-height: 5

library(fitdistrplus)

# Remove NAs and zeros for better fitting
fatalities_clean <- crashes$fatalities[!is.na(crashes$fatalities) & crashes$fatalities > 0]

# Fit distributions
fit_pois <- fitdist(fatalities_clean, "pois")
fit_nbinom <- fitdist(fatalities_clean, "nbinom")

# Create comparison plot
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
plot(fit_pois)
mtext("Poisson Fit", side = 3, line = -2, outer = TRUE, font = 2)

par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
plot(fit_nbinom)
mtext("Negative Binomial Fit", side = 3, line = -2, outer = TRUE, font = 2)
```

```{r}
#| label: aic-comparison

# Compare fits
tibble(
  Distribution = c("Poisson", "Negative Binomial"),
  AIC = c(fit_pois$aic, fit_nbinom$aic),
  BIC = c(fit_pois$bic, fit_nbinom$bic)
) %>%
  kable(digits = 2, caption = "Distribution fit comparison (lower is better)")
```

## Temporal Trends

```{r}
#| label: fig-temporal
#| fig-cap: "Temporal trends in aviation accidents"
#| fig-width: 10
#| fig-height: 6

p1 <- ggplot(crashes, aes(x = year, y = fatalities)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Fatalities Over Time",
       x = "Year", y = "Fatalities")

p2 <- crashes %>%
  group_by(year) %>%
  summarise(n_crashes = n(),
            total_fatalities = sum(fatalities, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = n_crashes)) +
  geom_line() +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Number of Crashes Over Time",
       x = "Year", y = "Number of Crashes")

p1 / p2
```

## Categorical Predictors

```{r}
#| label: fig-categorical
#| fig-cap: "Fatalities by aircraft type (top 15 types)"
#| fig-width: 10
#| fig-height: 6

crashes %>%
  filter(!is.na(aircraft_type) & aircraft_type != "") %>%
  group_by(aircraft_type) %>%
  summarise(
    mean_fatalities = mean(fatalities, na.rm = TRUE),
    median_fatalities = median(fatalities, na.rm = TRUE),
    n = n()
  ) %>%
  filter(n >= 10) %>%
  arrange(desc(mean_fatalities)) %>%
  head(15) %>%
  ggplot(aes(x = reorder(aircraft_type, mean_fatalities), 
             y = mean_fatalities)) +
  geom_point(aes(size = n), color = "steelblue") +
  geom_segment(aes(xend = aircraft_type, y = 0, yend = mean_fatalities),
               color = "steelblue") +
  coord_flip() +
  labs(title = "Mean Fatalities by Aircraft Type",
       subtitle = "Size indicates number of crashes",
       x = "Aircraft Type",
       y = "Mean Fatalities",
       size = "Number of\nCrashes")
```

## Hierarchical Structure Exploration

```{r}
#| label: fig-hierarchical
#| fig-cap: "Variation in fatalities across groups"
#| fig-width: 10
#| fig-height: 8

crashes %>%
  filter(!is.na(aircraft_type) & aircraft_type != "") %>%
  group_by(aircraft_type) %>%
  filter(n() >= 5) %>%
  ungroup() %>%
  ggplot(aes(x = reorder(aircraft_type, fatalities, FUN = median), 
             y = fatalities)) +
  geom_boxplot(fill = "steelblue", alpha = 0.5) +
  coord_flip() +
  labs(title = "Distribution of Fatalities by Aircraft Type",
       x = "Aircraft Type",
       y = "Fatalities")
```

The variation in fatality distributions across aircraft types suggests that a hierarchical model with partial pooling by aircraft type would be appropriate.

# Likelihood Selection

Based on the EDA:

1.  Response variable: Fatalities (count data)
2.  Distribution characteristics based on variance-to-mean ratio
3.  Recommended likelihood: Negative Binomial or Poisson
4.  Hierarchical grouping: Aircraft type, operator, or region

# Proposed Models

## Model 1: Non-hierarchical Model

```{r}
#| label: model1-description
#| eval: false

# Negative Binomial regression with year as predictor
# fatalities ~ year
# Non-hierarchical (pooled across all aircraft types)
```

## Model 2: Hierarchical Model

```{r}
#| label: model2-description
#| eval: false

# Negative Binomial regression with varying intercepts
# fatalities ~ year + (1 | aircraft_type)
# Hierarchical with partial pooling by aircraft type
```

# Next Steps

1.  Define informative priors based on domain knowledge
2.  Fit models using brms
3.  Conduct convergence diagnostics
4.  Perform posterior predictive checks
5.  Compare models using LOO-CV
6.  Sensitivity analysis

# References

Data source: planecrashinfo.com

# Appendix

## Session Info

```{r}
#| label: session-info

sessionInfo()
```