## Complete Pooling Model

With $i$ representing individual crashes, $n_i$ representing the number of people aboard, and $p_i$ representing the fatality probability of a crash, the model is:

$$\begin{align*}
\text{Likelihood:} \quad & \\
y_i &\sim \text{Binomial}(n_i, p_i) \\
\text{logit}(p_i) &= \alpha + \beta \cdot \text{year\_scaled}_i \\[1em]
\text{Priors:} \quad & \\
\alpha &\sim \mathcal{N}(1.4,\, 1) \\
\beta &\sim \mathcal{N}(0,\, 0.5)
\end{align*}$$

The complete pooling model is the simplest of our three models. The model assumes that each manufacturer shares the same intercept and linear coefficient. The weakness of this model is that it is not able to capture the potential differences between the manufacturers.

### Model

```{r}
#| label: complete_pool_model
#| cache: true

complete_pool_formula <- bf(
  fatalities | trials(aboard) ~ 1 + year_scaled,
  family = binomial(link = "logit")
  )

priors <- c(
  prior(normal(1.4, 1), class = Intercept),
  prior(normal(0, 0.5), class = b)
  )

complete_pool_model <- brm(
  formula = complete_pool_formula,
  prior = priors,
  data = crashes_filtered,
  seed = 42,
  refresh = 0,
  cores = 4,
  backend = "cmdstanr"
)
```

### Convergence Diagnostics

```{r}
#| label: complete_pool_model_summary

summary(complete_pool_model)
```

The $\hat{R}$ value for both the intercept term and the year coefficient is 1.00, indicating almost perfect chain overlapping. For both parameters, the bulk ESS values are over 3000 and the tail ESS values are over 2000. These values are great and definitely sufficient. The model had no divergences.

### Posterior Predictive Check

```{r}
#| label: complete_pool_model_fit_check
#| fig-width: 12
#| fig-height: 4

pp_check(complete_pool_model, ndraws = 100, type = "ecdf_overlay") +
  labs(title = "Complete Pool — ECDF overlay")

pp_check(complete_pool_model, ndraws = 100, type = "stat", stat = "mean") +
  labs(title = "Complete Pool — mean comparison")

pp_check(complete_pool_model, ndraws = 100, type = "stat", stat = "sd") +
  labs(title = "Complete Pool — SD comparison")
```

The complete pooling model ECDF follows the observed ECDF better at low fatalities than the no pooling model, but it still struggles in the 25–100 fatalities region.

The mean comparison plot shows that the observed mean is almost at the exact center of the model simulated fatalities distribution, meaning that this model also estimates the overall fatality rate correctly.

The standard deviation comparison plot shows that this model is also under-dispersed, but the effect is smaller than with the no pooling model.

### Predictions Over Time

```{r}
#| label: complete_pool_model_visualisation

pred_data <- tibble(
  year_scaled = seq(min(crashes_filtered$year_scaled),
                    max(crashes_filtered$year_scaled),
                    length.out = 100),
  aboard = round(mean(crashes_filtered$aboard))
)

predictions <- pred_data %>%
  add_epred_draws(complete_pool_model, ndraws = 1000) %>%
  mutate(.epred_rate = .epred / aboard)

ggplot() +
  stat_lineribbon(
    data = predictions,
    aes(x = year_scaled, y = .epred_rate),
    .width = c(0.50, 0.80, 0.95),
    alpha = 0.5
  ) +
  geom_point(
    data = crashes_filtered,
    aes(x = year_scaled, y = fatality_rate),
    alpha = 0.2,
    size = 1
  ) +
  scale_fill_brewer(palette = "Blues") +
  labs(
    title = "Complete Pooling Model: Fatality Rate Over Time",
    subtitle = paste0("Predictions for flights with ",
                     round(mean(crashes_filtered$aboard)), " people aboard (mean)"),
    x = "Year (Standardized)",
    y = "Fatality Rate",
    fill = "Credible Interval"
  )
```

### Prior Sensitivity Analysis

We conduct a prior sensitivity analysis on this model by trying out 3 different prior settings and comparing the estimated parameters. The priors we are using are:

- **Model prior:** Intercept prior: $\mathcal{N}(1.4, 1)$, year coefficient prior: $\mathcal{N}(0, 0.5)$ (same as the model).
- **Wide prior:** Intercept prior: $\mathcal{N}(1.4, 3)$, year coefficient prior: $\mathcal{N}(0, 1.5)$ (3 times as wide for both).
- **Shifted prior:** Intercept prior: $\mathcal{N}(3.4, 1)$, year coefficient prior: $\mathcal{N}(2, 0.5)$ (both priors shifted by 2 to the right).

| Prior Distribution | $\alpha$ | $\beta$ |
|---|---|---|
| Model prior | 0.81 | −0.18 |
| Wide prior | 0.81 | −0.18 |
| Shifted prior | 0.81 | −0.18 |

The parameters are identical for all three priors. From these results it can be concluded that the model is very insensitive to the prior choice.
