```{r}
#| label: setup
#| include: false

# List of required packages
required_packages <- c("tidyverse", "brms", "bayesplot", "loo", "posterior",
                       "skimr", "knitr", "patchwork", "ggdist", "tidybayes")

# Check which packages are not installed
missing_packages <- required_packages[!(required_packages %in% installed.packages()[,"Package"])]

# Install missing packages
if(length(missing_packages) > 0) {
  install.packages(missing_packages)
}

# Load all packages
for(pkg in required_packages) {
  library(pkg, character.only = TRUE)
}

# Setup cmdstan in home directory
# USED FOR JUPYTER HUB SETUP!
# cmdstan_path <- path.expand("~/.cmdstan/cmdstan-2.37.0")
# if(!dir.exists(cmdstan_path)) {
#   cmdstanr::install_cmdstan(dir = "~/.cmdstan")
# } else {
#   cmdstanr::set_cmdstan_path(cmdstan_path)
# }

# plotting theme
theme_set(theme_minimal())

# use parallel computing if possible
options(mc.cores = parallel::detectCores())
```

```{r}
#| label: load-data
#| include: false

# Load your plane crash data
crashes <- read_csv("data/aircrashdata.csv")

# Clean column names and filter valid observations
crashes <- crashes %>%
  rename(
    year = Year,
    quarter = Quarter,
    month = Month,
    day = Day,
    country = `Country/Region`,
    aircraft_manufacturer = `Aircraft Manufacturer`,
    aircraft_type = Aircraft,
    location = Location,
    operator = Operator,
    ground = Ground,
    fatalities = `Fatalities (air)`,
    aboard = Aboard
  ) %>%
  # Filter valid observations
  filter(
    aboard > 0,
    fatalities <= aboard,
  ) %>%
  mutate(
    # Calculate fatality rate once
    fatality_rate = fatalities / aboard,
    # Standardize year for better convergence
    year_scaled = scale(year)[,1],
  )

```

```{r}
#| label: data-summary
#| include: false

# overview statistics (numeric variables only)
crashes %>%
  select(where(is.numeric)) %>%
  skim()

# first rows of dataset
head(crashes) %>% 
  kable(caption = "First few rows of plane crash data")
```

## Crashes Over Time

```{r}
#| label: num-of-crashes
#| fig-cap: "Temporal trends in aviation accidents"
#| fig-width: 12
#| fig-height: 4

# Figure 1: Number of crashes - data
fig1 <- crashes %>%
  group_by(year) %>%
  summarise(n_crashes = n()) %>%
  ggplot(aes(x = year, y = n_crashes)) +
  geom_line() +
  labs(title = "Number of Crashes (Data)",
       x = "Year", y = "Number of Crashes")

# Figure 2: Number of crashes - trend
fig2 <- crashes %>%
  group_by(year) %>%
  summarise(n_crashes = n()) %>%
  ggplot(aes(x = year, y = n_crashes)) +
  geom_smooth(method = "loess", color = "red", se = TRUE) +
  labs(title = "Number of Crashes (Trend)",
       x = "Year", y = "Number of Crashes")

fig1 + fig2
```

## Fatalities Over Time

```{r}
#| label: fatalities-over-time
#| fig-cap: "Trends in aviation accidents"
#| fig-width: 12
#| fig-height: 4

# Figure 3: Fatalities - data points
fig3 <- crashes %>%
  ggplot(aes(x = year, y = fatalities)) +
  geom_point(alpha = 0.3) +
  labs(title = "Fatalities (Data Points)",
       x = "Year", y = "Fatalities")

# Figure 4: Fatalities - trend
fig4 <- crashes %>%
  ggplot(aes(x = year, y = fatalities)) +
  geom_smooth(method = "loess", color = "red", se = TRUE) +
  labs(title = "Fatalities (Trend)",
       x = "Year", y = "Fatalities")

fig3 + fig4
```

## Fatality Rate Over Time

```{r}
#| label: fatality-rate-over-time
#| fig-cap: "Mean fatality rate by year"
#| fig-width: 10
#| fig-height: 4

# Compute yearly mean fatality rate
rate_ts <- crashes %>%
  group_by(year) %>%
  summarise(
    mean_rate = mean(fatality_rate, na.rm = TRUE),
    sd_rate = sd(fatality_rate, na.rm = TRUE),
    n = n(),
    se = ifelse(n > 1, sd_rate / sqrt(n), NA_real_)
  )

# Figure 5: Fatality rate - data points
fig5 <- crashes %>%
  ggplot(aes(x = year, y = fatality_rate)) +
  geom_point(alpha = 0.3) +
  labs(title = "Fatality Rate (Data Points)",
       x = "Year", y = "Fatality Rate")

# Figure 6: Fatality rate - trend
fig6 <- ggplot(rate_ts, aes(x = year, y = mean_rate)) +
  geom_smooth(method = "loess", color = "red", se = TRUE) +
  labs(title = "Fatality Rate (Trend)",
       x = "Year", y = "Mean Fatality Rate")

fig5 + fig6
```

<!-- ```{r}
#| label: save-images
#| include: false

# Save all figures
ggsave("figures/crashes-data.pdf", fig1, width = 5, height = 3.5)
ggsave("figures/crashes-trend.pdf", fig2, width = 5, height = 3.5)
ggsave("figures/fatalities-data.pdf", fig3, width = 5, height = 3.5)
ggsave("figures/fatalities-trend.pdf", fig4, width = 5, height = 3.5)
ggsave("figures/fatality-rate-data.pdf", fig5, width = 5, height = 3.5)
ggsave("figures/fatality-rate-trend.pdf", fig6, width = 5, height = 3.5)
``` -->


To conduct a modern analysis on plane crash fatalities, we do not want to include observations from far back in the history. We need some reasonable cutoff point that keeps sufficient amount of data and allows us to model modern aircraft data. The fatalities trend shows a turning point around 1975, so it is reasonable to choose that as our observation cutoff year. This filtering brings our total observation count from 5013 down to 2188, so we are still left with plenty of data. We also only include observations from manufacturers with at least 10 crashes present in the data to stabilize the no pooling and partial pooling models. This brought our data down to 1766 observations.

1975 is a good cutoff date, because it resembles a period that shares lots of characteristics with modern flight travel. The 1950s and 1960s saw an upsurge in commercial flights, aircraft carriers, and international flying routes [@airwaysmag_commercial_flight]. This upsurge was accompanied by great technological innovations. The most notable ones were the jet engine in the 1950s (safer and more reliable than the earlier piston engines) and the introduction of electronic instruments to the cockpit in the 1970s [@allianz_aviation_safety].

```{r}
#| label: filtered-data-summary

len_before <- nrow(crashes)
crashes_filtered <- crashes %>% filter(year > 1975)
len_after <- nrow(crashes_filtered)

cat("Observations before filtering:", len_before, "\n")
cat("Observations after filtering:", len_after, "\n")
cat("Rows removed:", len_before - len_after, "\n")
```

```{r}
#| label: manufacturer-filtering

len_before <- nrow(crashes_filtered)
n_manufacturers_before <- n_distinct(crashes_filtered$aircraft_manufacturer)

min_crashes <- 10
crashes_filtered <- crashes_filtered %>%
  group_by(aircraft_manufacturer) %>%
  filter(n() >= min_crashes) %>%
  ungroup()

len_after <- nrow(crashes_filtered)
n_manufacturers_after <- n_distinct(crashes_filtered$aircraft_manufacturer)

cat("Observations before filtering:", len_before, "\n")
cat("Observations after filtering:", len_after, "\n")
cat("Rows removed:", len_before - len_after, "\n")
cat("Manufacturers before filtering:", n_manufacturers_before, "\n")
cat("Manufacturers after filtering:", n_manufacturers_after, "\n")
```

```{r}
#| label: filtered-data-numeric-summary

crashes_filtered %>%
  select(where(is.numeric)) %>%
  skim()
```

## Empirical Fatality Rates by Manufacturer

```{r}
#| label: model3-empirical-rates
#| fig-width: 10
#| fig-height: 8

# Compute empirical fatality rates by manufacturer (≥10 crashes)
empirical_rates <- crashes_filtered %>%
  group_by(aircraft_manufacturer) %>%
  summarise(
    mean_rate = mean(fatality_rate, na.rm = TRUE),
    sd_rate = sd(fatality_rate, na.rm = TRUE),
    n = n(),
    se = ifelse(n > 1, sd_rate / sqrt(n), NA_real_),
    .groups = "drop"
  ) %>%
  mutate(
    lower_ci = mean_rate - 1.96 * se,
    upper_ci = mean_rate + 1.96 * se
  ) %>%
  arrange(mean_rate)

# Plot empirical rates
empirical_rates %>%
  mutate(aircraft_manufacturer = fct_reorder(aircraft_manufacturer, mean_rate)) %>%
  ggplot(aes(x = mean_rate, y = aircraft_manufacturer)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci), height = 0.2) +
  geom_vline(
    xintercept = mean(crashes_filtered$fatality_rate), 
    linetype = "dashed", 
    color = "red"
  ) +
  labs(
    title = "Empirical Fatality Rates by Manufacturer",
    subtitle = "Data only (no pooling); dashed line = overall mean",
    x = "Fatality Rate", 
    y = "Manufacturer"
  ) +
  theme_minimal()
```