```{r}
#| label: setup
#| include: false

library(tidyverse)
library(brms)
library(bayesplot)
library(loo)
library(posterior)
library(skimr)
library(knitr)
library(patchwork)
library(ggdist)
library(tidybayes)

# plotting theme
theme_set(theme_minimal())

# use parallel computing if possible
options(mc.cores = parallel::detectCores())
```

```{r}
#| label: load-data
#| include: false

# Load your plane crash data
crashes <- read_csv("data/aircrashdata.csv")

# Clean column names and filter valid observations
crashes <- crashes %>%
  rename(
    year = Year,
    quarter = Quarter,
    month = Month,
    day = Day,
    country = `Country/Region`,
    aircraft_manufacturer = `Aircraft Manufacturer`,
    aircraft_type = Aircraft,
    location = Location,
    operator = Operator,
    ground = Ground,
    fatalities = `Fatalities (air)`,
    aboard = Aboard
  ) %>%
  # Filter valid observations
  filter(
    aboard > 0,
    fatalities <= aboard,
  ) %>%
  mutate(
    # Calculate fatality rate once
    fatality_rate = fatalities / aboard,
    # Standardize year for better convergence
    year_scaled = scale(year)[,1],
  )

```

```{r}
#| label: data-summary
#| include: false

# overview statistics (numeric variables only)
crashes %>%
  select(where(is.numeric)) %>%
  skim()

# first rows of dataset
head(crashes) %>% 
  kable(caption = "First few rows of plane crash data")
```

## Crashes Over Time

```{r}
#| label: num-of-crashes
#| fig-cap: "Temporal trends in aviation accidents"
#| fig-width: 12
#| fig-height: 4

# Plot with just the line (data points)
p1 <- crashes %>%
  group_by(year) %>%
  summarise(n_crashes = n(),
            total_fatalities = sum(fatalities, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = n_crashes)) +
  geom_line() +
  labs(title = "Number of Crashes Over Time (Data)",
       x = "Year", y = "Number of Crashes")

# Plot with just the trend line
p2 <- crashes %>%
  group_by(year) %>%
  summarise(n_crashes = n(),
            total_fatalities = sum(fatalities, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = n_crashes)) +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Number of Crashes Over Time (Trend)",
       x = "Year", y = "Number of Crashes")

p1 + p2
```

## Fatalities Over Time

```{r}
#| label: fatalities-over-time
#| fig-cap: "Trends in aviation accidents"
#| fig-width: 12
#| fig-height: 4

# Plot with just data points
p1 <- crashes %>%
  ggplot(aes(x = year, y = fatalities)) +
  geom_point(alpha = 0.3) +
  labs(title = "Fatalities Over Time (Data Points)",
       x = "Year", y = "Fatalities")

# Plot with just the trend line
p2 <- crashes %>%
  ggplot(aes(x = year, y = fatalities)) +
  geom_smooth(method = "loess", color = "red") +
  labs(title = "Fatalities Over Time (Trend)",
       x = "Year", y = "Fatalities")

p1 + p2
```

## Fatality Rate Over Time

```{r}
#| label: fatality-rate-over-time
#| fig-cap: "Mean fatality rate by year"
#| fig-width: 10
#| fig-height: 4

# Compute yearly mean fatality rate with standard error
rate_ts <- crashes %>%
  group_by(year) %>%
  summarise(
    mean_rate = mean(fatality_rate, na.rm = TRUE),
    sd_rate = sd(fatality_rate, na.rm = TRUE),
    n = n(),
    se = ifelse(n > 1, sd_rate / sqrt(n), NA_real_)
  )

# Plot 1: Individual data points
p1 <- crashes %>%
  ggplot(aes(x = year, y = fatality_rate)) +
  geom_point(alpha = 0.3) +
  labs(title = "Fatality Rate Over Time (Data Points)",
       x = "Year", y = "Fatality Rate")


# Plot 2: Trend
p2 <- ggplot(rate_ts, aes(x = year, y = mean_rate)) +
  geom_smooth(method = "loess", color = "red", se = TRUE) +
  labs(title = "Fatality Rate Trend",
       x = "Year", y = "Mean Fatality Rate")

# Combine plots side by side
p1 + p2
```

To conduct a modern analysis on plane crash fatalities, we do not want to include observations from far back in the history. We need some reasonable cutoff point that keeps sufficient amount of data and allows us to model modern aircraft data. The fatalities trend shows a turning point at around 1975 so it is reasonable to choose that as our observation cutoff year. 
```{r}
len_before <- nrow(crashes)
crashes <- crashes %>% filter(year > 1975)
len_after <- nrow(crashes)

cat("Observations before filtering:", len_before, "\n")
cat("Observations after filtering:", len_after, "\n")
cat("Rows removed:", len_before - len_after, "\n")
```

```{r}
crashes %>%
  select(where(is.numeric)) %>%
  skim()
```

## Empirical Fatality Rates by Manufacturer

```{r}
#| label: model3-empirical-rates
#| fig-width: 10
#| fig-height: 8
min_crashes <- 10
crashes_filtered <- crashes %>%
  group_by(aircraft_manufacturer) %>%
  filter(n() >= min_crashes) %>%
  ungroup()
# Compute empirical fatality rates by manufacturer (≥10 crashes)
empirical_rates <- crashes_filtered %>%
  group_by(aircraft_manufacturer) %>%
  summarise(
    mean_rate = mean(fatality_rate, na.rm = TRUE),
    sd_rate = sd(fatality_rate, na.rm = TRUE),
    n = n(),
    se = ifelse(n > 1, sd_rate / sqrt(n), NA_real_),
    .groups = "drop"
  ) %>%
  mutate(
    lower_ci = mean_rate - 1.96 * se,
    upper_ci = mean_rate + 1.96 * se
  ) %>%
  arrange(mean_rate)

# Plot empirical rates
empirical_rates %>%
  mutate(aircraft_manufacturer = fct_reorder(aircraft_manufacturer, mean_rate)) %>%
  ggplot(aes(x = mean_rate, y = aircraft_manufacturer)) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci), height = 0.2) +
  geom_vline(xintercept = mean(crashes_filtered$fatality_rate), linetype = "dashed", color = "red") +
  labs(title = "Empirical Fatality Rates by Manufacturer",
       subtitle = "Data only (no pooling); dashed line = overall mean",
       x = "Fatality Rate", 
       y = "Manufacturer") +
  theme_minimal()
```