## Complete Pooling Model

This model treats all plane crashes as coming from a single population, ignoring differences between countries, manufacturers, or time periods.

### Fit Complete Pooling Model

We'll model the fatality rate using a binomial model with logit link. This approach:
- Naturally handles proportions including 0s and 1s
- Models fatalities out of total aboard (respects the actual counts)
- Uses the logit transformation internally
- More appropriate for count data than beta regression

```{r}
#| label: model2
#| cache: true

# Check the data structure
crashes %>%
  summarise(
    n_zeros = sum(fatalities == 0),
    n_all_died = sum(fatalities == aboard),
    min_rate = min(fatality_rate),
    max_rate = max(fatality_rate)
  )

# Fit complete pooling model with binomial family
complete_pool_model <- brm(
  fatalities | trials(aboard) ~ 1 + year_scaled,
  data = crashes,
  family = binomial(link = "logit"),
  prior = c(
    prior(normal(1.4, 1), class = Intercept),
    prior(normal(0, 0.5), class = b)
  ),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  cores = 4,
  seed = 123,
  backend = "cmdstanr",
  refresh = 0
)
```

### Model Summary

```{r}
#| label: model2 summary

summary(complete_pool_model)
```

### Posterior Predictive Checks

```{r}
#| label: model2 fit check

# Check model fit
pp_check(complete_pool_model, ndraws = 100)
```

### Visualize Predictions

```{r}
#| label: model2 visualisation

# Create prediction data
# For binomial models, we need to specify the number of trials
# We'll use the mean number aboard as a reference point
pred_data <- tibble(
  year_scaled = seq(min(crashes$year_scaled), 
                    max(crashes$year_scaled), 
                    length.out = 100),
  aboard = round(mean(crashes$aboard))  # Use mean as reference
)

# Generate predictions (these are expected counts of fatalities)
predictions <- pred_data %>%
  add_epred_draws(complete_pool_model, ndraws = 1000) %>%
  mutate(.epred_rate = .epred / aboard)  # Convert to rate for plotting

# Plot predictions with uncertainty
ggplot() +
  stat_lineribbon(
    data = predictions,
    aes(x = year_scaled, y = .epred_rate),
    .width = c(0.50, 0.80, 0.95),
    alpha = 0.5
  ) +
  geom_point(
    data = crashes,
    aes(x = year_scaled, y = fatality_rate),
    alpha = 0.2,
    size = 1
  ) +
  scale_fill_brewer(palette = "Blues") +
  labs(
    title = "Complete Pooling Model: Fatality Rate Over Time",
    subtitle = paste0("Predictions shown for flights with ", 
                     round(mean(crashes$aboard)), " people aboard (mean)"),
    x = "Year (Standardized)",
    y = "Fatality Rate",
    fill = "Credible Interval"
  ) +
  theme_minimal()
```

### Extract and Interpret Coefficients

```{r}
#| label: model2 interpretation

# Extract posterior samples
posterior_samples <- as_draws_df(complete_pool_model)

# Summarize key parameters on logit scale
posterior_samples %>%
  select(b_Intercept, b_year_scaled) %>%
  pivot_longer(everything(), names_to = "parameter", values_to = "value") %>%
  group_by(parameter) %>%
  summarise(
    mean = mean(value),
    sd = sd(value),
    q025 = quantile(value, 0.025),
    q975 = quantile(value, 0.975)
  )

# Convert intercept to probability scale for interpretation
posterior_samples %>%
  mutate(baseline_prob = plogis(b_Intercept)) %>%
  summarise(
    mean_baseline_fatality_rate = mean(baseline_prob),
    q025 = quantile(baseline_prob, 0.025),
    q975 = quantile(baseline_prob, 0.975)
  )
```

### Model Interpretation

The complete pooling model assumes:
- All crashes share the same underlying fatality rate pattern
- Temporal trend captured by `year_scaled` coefficient
- No variation between countries, manufacturers, or other grouping factors
- This serves as a baseline for comparison with partial pooling (hierarchical) models