## No Pooling Model

This model treats each aircraft manufacturer completely separately, fitting independent fatality rate patterns for each manufacturer without borrowing any information between groups.

### Fit No Pooling Model

We'll fit separate binomial models per manufacturer using fixed effects. This approach:
- Gives each manufacturer its own intercept and temporal trend
- Uses no information sharing (no pooling) between manufacturers
- Requires sufficient data per manufacturer (minimum 10 crashes)
- Uses the same binomial family and logit link as other models

```{r}
#| label: no_pool_model
#| cache: true

# Check the data structure
crashes %>%
  summarise(
    n_manufacturers = n_distinct(aircraft_manufacturer),
    n_crashes = n(),
    n_zeros = sum(fatalities == 0),
    n_all_died = sum(fatalities == aboard),
    min_rate = min(fatality_rate),
    max_rate = max(fatality_rate)
  )

# Fit no-pooling model with separate intercept and slope per manufacturer
# Using 0 to remove the global intercept, then add manufacturer-specific intercepts
no_pool_model <- brm(
  fatalities | trials(aboard) ~ 0 + aircraft_manufacturer + aircraft_manufacturer:year_scaled,
  data = crashes,
  family = binomial(link = "logit"),
  prior = c(
    prior(normal(0, 1), class = b)
  ),
  chains = 4,
  iter = 2000,
  warmup = 1000,
  cores = 4,
  seed = 123,
  backend = "rstan",
  refresh = 0
)
```

### Model Summary

```{r}
#| label: no_pool_model summary

summary(no_pool_model)
```

### Posterior Predictive Checks

```{r}
#| label: no_pool_model fit check

# Check model fit
pp_check(no_pool_model, ndraws = 100)
```

### Visualize Predictions

```{r}
#| label: no_pool_model visualisation
#| fig-width: 14
#| fig-height: 10

# Select top manufacturers for visualization
top_manufacturers <- crashes %>%
  count(aircraft_manufacturer, sort = TRUE) %>%
  slice_head(n = 9) %>%
  pull(aircraft_manufacturer)

# Create prediction data for these manufacturers
pred_data <- expand_grid(
  aircraft_manufacturer = top_manufacturers,
  year_scaled = seq(min(crashes$year_scaled), 
                    max(crashes$year_scaled), 
                    length.out = 100)
) %>%
  mutate(aboard = round(mean(crashes$aboard)))

# Generate predictions (these are expected counts of fatalities)
predictions <- pred_data %>%
  add_epred_draws(no_pool_model, ndraws = 1000) %>%
  mutate(.epred_rate = .epred / aboard)  # Convert to rate for plotting

# Plot predictions with uncertainty, faceted by manufacturer
ggplot() +
  stat_lineribbon(
    data = predictions,
    aes(x = year_scaled, y = .epred_rate),
    .width = c(0.50, 0.80, 0.95),
    alpha = 0.5
  ) +
  geom_point(
    data = crashes %>% filter(aircraft_manufacturer %in% top_manufacturers),
    aes(x = year_scaled, y = fatality_rate),
    alpha = 0.2,
    size = 1
  ) +
  facet_wrap(~ aircraft_manufacturer, scales = "free_y", ncol = 3) +
  scale_fill_brewer(palette = "Blues") +
  labs(
    title = "No Pooling Model: Manufacturer-Specific Fatality Rates",
    subtitle = "Each manufacturer has independent intercept and slope (top 9 by crash count)",
    x = "Year (Standardized)",
    y = "Fatality Rate",
    fill = "Credible Interval"
  ) +
  theme_minimal()
```

### Extract and Interpret Coefficients

```{r}
#| label: no_pool_model interpretation

# Extract posterior samples
posterior_samples <- as_draws_df(no_pool_model)

# Get coefficient names
coef_names <- colnames(posterior_samples)[grep("^b_", colnames(posterior_samples))]

# Summarize all coefficients
posterior_samples %>%
  select(all_of(coef_names)) %>%
  pivot_longer(everything(), names_to = "parameter", values_to = "value") %>%
  group_by(parameter) %>%
  summarise(
    mean = mean(value),
    sd = sd(value),
    q025 = quantile(value, 0.025),
    q975 = quantile(value, 0.975),
    .groups = "drop"
  ) %>%
  arrange(mean)

# Convert manufacturer intercepts to probability scale for interpretation
# Extract just the main effects (intercepts) for top manufacturers
manuf_intercepts <- posterior_samples %>%
  select(starts_with("b_aircraft_manufacturer")) %>%
  select(!ends_with("year_scaled")) %>%
  pivot_longer(everything(), names_to = "manufacturer", values_to = "logit_value") %>%
  mutate(
    manufacturer = str_remove(manufacturer, "^b_aircraft_manufacturer"),
    probability = plogis(logit_value)
  ) %>%
  group_by(manufacturer) %>%
  summarise(
    mean_fatality_rate = mean(probability),
    q025 = quantile(probability, 0.025),
    q975 = quantile(probability, 0.975),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_fatality_rate))

print(manuf_intercepts)
```

### Model Interpretation

The no-pooling model assumes:
- Each manufacturer has its own independent fatality rate pattern
- No information is shared between manufacturers (no pooling)
- Temporal trends vary by manufacturer
- This provides maximum flexibility but requires sufficient data per manufacturer
- Comparison point: shows what the data suggests when we allow complete independence
