## No Pooling Model

This model treats every airplane manufacturer as their own independent entity, without sharing any information across any other possible grouping in the data. With $i$ being the index of a plane crash and $m[i]$ being the manufacturer $m$ for crash $i$:

$$\begin{align*}
\text{Likelihood:} \quad & \\
y_i &\sim \text{Binomial}(n_i, p_i) \\
\text{logit}(p_i) &= \alpha_{m[i]} + \beta_{m[i]} \cdot \text{year\_scaled}_i \\[1em]
\text{Priors:} \quad & \\
\alpha_m &\sim \mathcal{N}(0,\, 1) \quad \text{for } m = 1, \ldots, M \\
\beta_m &\sim \mathcal{N}(0,\, 1) \quad \text{for } m = 1, \ldots, M
\end{align*}$$

This means that the model assumes that each crash has its own fatality rate based on the manufacturer and the year, and crashes from other manufacturers do not influence each other in any way. With this model, the estimates for manufacturers with only a few crashes in the data become unstable, because they can't borrow information from other manufacturers.

### Model

```{r}
#| label: no_pool_model
#| cache: true

no_pool_formula <- bf(
	fatalities | trials(aboard) ~ 0 + aircraft_manufacturer + aircraft_manufacturer:year_scaled,
	family = binomial(link = "logit")
	)

priors <- c(
		prior(normal(0, 1), class = b)
		)

no_pool_model <- brm(
	formula = no_pool_formula,
	prior = priors,
  data = crashes_filtered,
	seed = 42,
	refresh = 0,
	cores = 4,
	backend = "cmdstanr"
)
```

### Convergence Diagnostics

```{r}
#| label: no_pool_model_summary

summary(no_pool_model)
```

The $\hat{R}$ value for each of the parameters is 1.00, which means all coefficients converged properly. The bulk ESS values vary on the scale [4200, 5800] and the tail ESS values on the scale [2600, 3500]. These ESS values are far more than is needed for a reliable posterior estimation, so no problems there. The model had no divergences.

### Posterior Predictive Check

```{r}
#| label: no_pool_model_fit_check
#| fig-width: 12
#| fig-height: 4

pp_check(no_pool_model, ndraws = 100, type = "ecdf_overlay") +
  labs(title = "No Pool — ECDF overlay")

pp_check(no_pool_model, ndraws = 100, type = "stat", stat = "mean") +
  labs(title = "No Pool — mean comparison")

pp_check(no_pool_model, ndraws = 100, type = "stat", stat = "sd") +
  labs(title = "No Pool — SD comparison")
```

The ECDF (empirical cumulative distribution function) shows that the model is slightly off when compared to the real data. We can see that at the start, the observed ECDF rises faster and higher than the model ECDF, indicating that the model predicts less crashes with close to 0 fatalities. However, in the 25–100 fatalities area, the observed ECDF lags behind the model ECDF meaning that the model estimates more crashes that lead to those fatality numbers.

The mean comparison plot shows that the true fatalities mean lies close to the center of the model simulated fatalities distribution, meaning that the model estimates the overall mean fatality rate correctly.

The standard deviation comparison plot shows that the true fatalities standard deviation is much larger than what the model estimates. In other words, the model predicts too little variability and is under-dispersed.

### Predictions by Manufacturer

```{r}
#| label: no_pool_model_visualisation
#| fig-width: 14
#| fig-height: 10

top_manufacturers <- crashes_filtered %>%
  count(aircraft_manufacturer, sort = TRUE) %>%
  slice_head(n = 9) %>%
  pull(aircraft_manufacturer)

pred_data <- expand_grid(
  aircraft_manufacturer = top_manufacturers,
  year_scaled = seq(min(crashes_filtered$year_scaled),
                    max(crashes_filtered$year_scaled),
                    length.out = 100)
) %>%
  mutate(aboard = round(mean(crashes_filtered$aboard)))

predictions <- pred_data %>%
  add_epred_draws(no_pool_model, ndraws = 1000) %>%
  mutate(.epred_rate = .epred / aboard)

ggplot() +
  stat_lineribbon(
    data = predictions,
    aes(x = year_scaled, y = .epred_rate),
    .width = c(0.50, 0.80, 0.95),
    alpha = 0.5
  ) +
  geom_point(
    data = crashes_filtered %>% filter(aircraft_manufacturer %in% top_manufacturers),
    aes(x = year_scaled, y = fatality_rate),
    alpha = 0.2,
    size = 1
  ) +
  facet_wrap(~ aircraft_manufacturer, scales = "free_y", ncol = 3) +
  scale_fill_brewer(palette = "Blues") +
  labs(
    title = "No Pooling Model: Manufacturer-Specific Fatality Rates",
    subtitle = "Each manufacturer has independent intercept and slope (top 9 by crash count)",
    x = "Year (Standardized)",
    y = "Fatality Rate",
    fill = "Credible Interval"
  )
```

### Prior Sensitivity Analysis

We conduct a prior sensitivity analysis on this model by trying out 3 different priors and comparing the estimated parameters of 4 manufacturers. The priors we are using are:

- **Model prior:** $\mathcal{N}(0, 1)$ (same as the model).
- **Wide prior:** $\mathcal{N}(0, 3)$ (a much wider distribution).
- **Shifted prior:** $\mathcal{N}(2, 1)$ (shifted to the right).

| Prior Distribution | Manufacturer | $\alpha$ | $\beta$ |
|---|---|---|---|
| Model $\mathcal{N}(0, 1)$ | Boeing | 0.72 | −0.34 |
| | Bell | 1.94 | 0.82 |
| | Lockheed | 0.16 | 1.00 |
| | Mil | 4.34 | −1.67 |
| Wide $\mathcal{N}(0, 3)$ | Boeing | 0.72 | −0.34 |
| | Bell | 2.20 | 0.63 |
| | Lockheed | 0.16 | 1.00 |
| | Mil | 5.16 | −2.30 |
| Shifted $\mathcal{N}(2, 1)$ | Boeing | 0.72 | −0.34 |
| | Bell | 1.96 | 0.96 |
| | Lockheed | 0.16 | 1.00 |
| | Mil | 4.40 | −1.70 |

We can see that the parameters do not vary much. For manufacturers with many observations in the data (Boeing: 248 observations, Lockheed: 106 observations), the parameters are identical for every prior. For manufacturers with a fewer number of observations (Mil: 41 observations, Bell: 33 observations), the parameter values show more variance:

$$\text{Mil: } \operatorname{Var}(\alpha) \approx 0.21,\quad \operatorname{Var}(\beta) \approx 0.11$$
$$\text{Bell: } \operatorname{Var}(\alpha) \approx 0.02,\quad \operatorname{Var}(\beta) \approx 0.03$$

These variances are tolerable and still show the same trend, just in different strengths. The results show that the model is stable under reasonable prior changes, and the main conclusions drawn from the model do not depend on the chosen prior.
